---
title: "Assignment 3"
---

This assignment demonstrates how to collect, map, and interpret US Census data using R. We use the `tidycensus`, `tigris`, `sf`, and `ggplot2` packages. The focus is on median household income and poverty counts by **county** in **California** using the **2023 ACS 5-year estimates**.

```{r, warning=FALSE, message=FALSE, results='hide', include=TRUE}
library(tidycensus)
library(tigris)
library(sf)
library(dplyr)
library(ggplot2)
library(readr)

options(tigris_use_cache = TRUE)
```

```{r, message=FALSE}
# 1) API key (make sure you have your Census API key!)
census_api_key("1679bfcd12ef3ba5b6103471f96325933d49aa1f", install = TRUE, overwrite = TRUE)
```

```{r}
# 2) Explore variables
vars <- load_variables(2023, "acs5", cache = TRUE)
vars |> dplyr::filter(grepl("^B19", name)) |> dplyr::slice_head(n = 10)
```

```{r}
# 3) Parameters (California)
state_abbr <- "CA"
geo_level  <- "county"   # options: state, county, tract, block group
my_vars    <- c(income = "B19013_001", poverty = "B17001_002")
year_acs   <- 2023
survey     <- "acs5"
```

```{r}
# 4) Download ACS data with geometry
acs <- get_acs(
  geography = geo_level,
  variables = my_vars,
  state = state_abbr,
  year = year_acs,
  survey = survey,
  geometry = TRUE
)
```

```{r}
# 5) Convert to wide format
acs_wide <- acs |>
  tidyr::pivot_wider(
    id_cols = c(GEOID, NAME, geometry),
    names_from = variable,
    values_from = c(estimate, moe)
  )
```

```{r, warning=FALSE}
# 6) Map median income
ggplot(acs_wide) +
  geom_sf(aes(fill = estimate_income), color = "gray", size = 0.1) +
  scale_fill_viridis_c(name = "Median HH Income") +
  labs(title = paste0("ACS ", year_acs, " 5-year: Median Income — ", state_abbr, " (", geo_level, ")"),
       caption = "Source: U.S. Census Bureau via tidycensus") +
  theme_minimal() + 
  theme(
  legend.position = c(0.99, 0.99),
  legend.justification = c("right", "top"),
  legend.title = element_text(size = 11),
  legend.text = element_text(size = 9),
  legend.background = element_rect(fill = alpha('white', 0.6), color = NA),
  plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
  plot.caption = element_text(hjust = 0.5, size = 9)
)
```

The choropleth map above visualizes **median household income** across California counties using data from the **2023 ACS 5-year estimates**. Counties are shaded using a viridis color scale, where **yellow indicates the highest income levels** and **dark purple indicates the lowest**.

From the map, it is clear that the **San Francisco Bay Area** (including counties such as **Santa Clara**, **San Mateo**, and **Marin**) has the **highest median household incomes**, with some exceeding **$150,000**. These counties stand out in bright yellow. On the other hand, **inland and northern counties** such as those in the **Central Valley** or far **Northern California** tend to have **lower median incomes**, frequently under **$75,000**.


```{r}
# 7) Top and bottom counties by poverty count
top10 <- acs_wide |>
  arrange(desc(estimate_poverty)) |>
  select(NAME, estimate_poverty, moe_poverty) |>
  slice_head(n = 10)

bottom10 <- acs_wide |>
  arrange(estimate_poverty) |>
  select(NAME, estimate_poverty, moe_poverty) |>
  slice_head(n = 10)

top10
bottom10
```

```{r}
# 8) Save the data as CSV (optional)
readr::write_csv(st_drop_geometry(acs_wide), "acs_data.csv")
```

